Analysis_Report

### 2019-04-11 analysis/analysis_report.rmd
first saccade direction analysis

@@

@@

>>>>

singleton_absent_omissions <- exptdata_means_byparticipant %>%
  ungroup() %>%
  filter(condition == "Singleton Search") %>%
  filter(dType == "Absent") %>%
  select(m) %>%
  pull()

@@

feature_absent_omissions <- exptdata_means_byparticipant %>%
  ungroup() %>%
  filter(condition == "Feature Search") %>%
  filter(dType == "Absent") %>%
  select(m) %>%
  pull()

@@

### Effect of Physical Salience
Effect of physical salience ANOVA:
```{r Effect of Physical Salience - Omissions}
omissions_salience.ANOVA <- aov_car(m ~ condition*dType + Error(sub/dType), data = filter(exptdata_means_byparticipant, dType != "High"), anova_table = list(es="pes"))

omissions_salience.ANOVA
```

Planned t-tests:
```{r Effect of Physical Salience - Omissions - Planned t-tests}
singleton_physicalsalience_omissions.ttest <- t.test(singleton_low_omissions, singleton_absent_omissions, paired = T)

singleton_physicalsalience_omissions.es <- dz_calculator(singleton_low_omissions, singleton_absent_omissions)

feature_physicalsalience_omissions.ttest <- t.test(feature_absent_omissions, feature_low_omissions, paired = T) # reverse the order here so can get positive t value

feature_physicalsalience_omissions.es <- dz_calculator(feature_low_omissions, feature_absent_omissions)
```

# Gaze Data Analyses

```{r Load Gaze Data, include=FALSE}
saccade_files <- dir(path = here::here("analysis", "data", "derived_data","ProcessedSaccadeData"), pattern = "^S")

saccade_filedata <- here::here("analysis", "data", "derived_data","ProcessedSaccadeData", saccade_files) %>%
  map(readMat) %>%
  map("saccadeSessionData") %>%
  map(1) %>%
  map(2) %>%
  modify(as.data.frame)
```

```{r Tidy Gaze Data, include = FALSE}
# collapse the data into a dataframe format
saccade_filedata <- saccade_filedata %>%
  map_df(bind_rows)

# give the columns appropriate headers
colnames(saccade_filedata)[1:13] <- c("subj", "trial", "latency", paste("loc", c(1:5), sep = "_"), "discardTrial", "anticipatorySaccade", "outsideFixation", "noSaccades", "noValidData")

discardSaccadeCounter <- saccade_filedata %>%
  mutate(bad_data_discard = case_when(noSaccades == 1 ~ 1,
                                      noValidData == 1 ~ 1,
                                      TRUE ~ 0)) %>%
  group_by(subj) %>%
  summarise(prop_discarded = mean(discardTrial)*100, prop_anticipatory = mean(anticipatorySaccade)*100, prop_outside_fix = mean(outsideFixation)*100, prop_bad_data = mean(bad_data_discard)*100)

totalSaccadeTrialsDiscarded <- discardSaccadeCounter %>%
  summarise_all(mean) %>%
  summarise_all(round, 1)

saccade_filedata <- saccade_filedata %>%
  filter(discardTrial != 1) %>% # discard trials marked to be discarded
  mutate(first_saccade_loc = loc_1*1+loc_2*2+loc_3*3+loc_4*4+loc_5*5)

saccadedata <- left_join(exptdata, saccade_filedata, by = c("sub" = "subj", "trial" = "trial")) %>% #join saccade data up with behavioural data
  mutate(saccade_to_target = if_else(targetLoc == first_saccade_loc, 1, 0)) %>% # create variables indicating whether first saccade went towards the target...
  mutate(saccade_to_singleton = if_else(dType == "Absent", 2, 
                                        if_else(distractLoc == first_saccade_loc, 1, 0))) %>% # singleton (for all except distractor absent trials)
  mutate(saccade_to_singleton = replace(saccade_to_singleton, saccade_to_singleton == 2, NA)) %>%
  mutate(saccade_to_nonsingleton = if_else((dType != "Absent" & targetLoc != first_saccade_loc & distractLoc != first_saccade_loc & first_saccade_loc < 5) | (dType == "Absent" & targetLoc != first_saccade_loc & first_saccade_loc < 5), 1, 0)) %>% # or non-singleton
  mutate(num_nonsingletons = if_else(dType == "Absent", 3, 2)) # add the number of nonsingletons for each trial type
```

## First saccade direction
```{r First saccade direction - means}
# calculate summary statistics for each participant
saccadeData_means_byparticipant <- saccadedata %>%
  ungroup() %>%
  mutate(sub = as.factor(sub)) %>%
  group_by(sub, condition, dType) %>%
  summarise(saccade_to_target = mean(saccade_to_target, na.rm = T),
            saccade_to_singleton = mean(saccade_to_singleton, na.rm = T),
            saccade_to_nonsingleton = mean(saccade_to_nonsingleton, na.rm = T)/mean(num_nonsingletons, na.rm = T),
            num_nonsingletons = mean(num_nonsingletons, na.rm = T))

targetmeans <- summarySEwithin2(data = saccadeData_means_byparticipant, measurevar = "saccade_to_target", betweenvars = "condition", withinvars = c("dType"), idvar = "sub", na.rm = T) %>%
  rename(proportionSaccades = saccade_to_target) %>%
  mutate(stimType = "Target") %>%
  select(-ends_with("Normed"))

singletonmeans <- summarySEwithin2(data = saccadeData_means_byparticipant, measurevar = "saccade_to_singleton", betweenvars = "condition", withinvars = c("dType"), idvar = "sub", na.rm = T) %>%
  rename(proportionSaccades = saccade_to_singleton) %>%
  mutate(stimType = "Singleton")%>%
  select(-ends_with("Normed"))

nonsingletonmeans <- summarySEwithin2(data = saccadeData_means_byparticipant, measurevar = "saccade_to_nonsingleton", betweenvars = "condition", withinvars = c("dType"), idvar = "sub", na.rm = T) %>%
  rename(proportionSaccades = saccade_to_nonsingleton) %>%
  mutate(stimType = "Nonsingleton")%>%
  select(-ends_with("Normed"))

# collapse together
saccadeData_means_bycondition <- bind_rows(targetmeans, singletonmeans, nonsingletonmeans) %>%
  mutate(stimType = factor(stimType, levels = c("Target","Nonsingleton","Singleton")))

saccadePlotDataParticipants <- gather(saccadeData_means_byparticipant, key = "stimType", value = "proportionSaccades", starts_with("saccade_to")) %>%
  mutate(stimType = case_when(stimType == "saccade_to_target" ~ "Target",
                              stimType == "saccade_to_nonsingleton" ~ "Nonsingleton",
                              stimType == "saccade_to_singleton" ~ "Singleton",
                              TRUE ~ ""))

captureData_means_byparticipant <- saccadeData_means_byparticipant %>%
  filter(dType != "Absent") %>%
  mutate(Capture = saccade_to_singleton - saccade_to_nonsingleton)

captureMeans <- summarySEwithin2(captureData_means_byparticipant, measurevar = "Capture", betweenvars = "condition", withinvars = "dType", idvar = "sub") %>%
  select(-ends_with("Normed"))
```

```{r First Saccade Direction - Plot, echo = FALSE}
saccadePlot_singleton <- saccadeData_means_bycondition %>%
  filter(condition == "Singleton Search", dType != "Absent") %>%
  ggplot(aes(group = dType, y = proportionSaccades, fill = dType, x = stimType, shape = dType, colour = dType)) +
  geom_line(aes(linetype = dType, colour = dType), na.rm = T, size = .5, position = position_dodge(width = .6)) +
  geom_point(aes(group = dType, y = proportionSaccades, x = stimType), na.rm = T, size = 1.5, alpha = .3, stroke = F, data = filter(saccadePlotDataParticipants, condition == "Singleton Search", dType != "Absent"), position = position_jitterdodge(dodge.width = .6, jitter.width = .2)) +
  geom_errorbar(aes(ymin = proportionSaccades-se, ymax = proportionSaccades+se), na.rm = T, linetype = "solid", position = position_dodge(width = .6, preserve = "total"), show.legend = FALSE, colour = "black", width = .3) +
  geom_point(na.rm = T, size = 3, position = position_dodge(width = .6), colour = "black") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = c(0,0), limits = c(0,.8)) +
  scale_linetype_manual(name = "Trial Type", values = c("solid", "longdash", "dotdash"), guide = guide_legend(), labels = c("High-reward", "Low-reward")) +
  scale_fill_manual(name = "Trial Type", values = vmacCols, labels = c("High-reward", "Low-reward")) +
  scale_colour_manual(name = "Trial Type", values = vmacCols, labels = c("High-reward", "Low-reward")) +
  scale_shape_manual(name = "Trial Type", guide = guide_legend(), values = c(21,22,23), labels = c("High-reward", "Low-reward")) +
  ylab("% First Saccades") + xlab("Stimulus Type") + labs(tag = "A", title = "Singleton Search") +
  my_theme

saccadePlot_feature <- saccadeData_means_bycondition %>%
  filter(condition == "Feature Search", dType != "Absent") %>%
  ggplot(aes(group = dType, y = proportionSaccades, fill = dType, x = stimType, shape = dType, colour = dType)) +
  geom_line(aes(linetype = dType, colour = dType), na.rm = T, size = .5, position = position_dodge(width = .6)) +
  geom_point(aes(group = dType, y = proportionSaccades, x = stimType), na.rm = T, size = 1.5, alpha = .3, stroke = F, data = filter(saccadePlotDataParticipants, condition == "Feature Search", dType != "Absent"), position = position_jitterdodge(dodge.width = .6, jitter.width = .2)) +
  geom_errorbar(aes(ymin = proportionSaccades-se, ymax = proportionSaccades+se), na.rm = T, linetype = "solid", position = position_dodge(width = .6, preserve = "total"), show.legend = FALSE, colour = "black", width = .3) +
  geom_point(na.rm = T, size = 3, position = position_dodge(width = .6), colour = "black") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = c(0,0), limits = c(0,.8)) +
  scale_linetype_manual(name = "Distractor Type", values = c("solid", "longdash", "dotdash"), guide = guide_legend()) +
  scale_fill_manual(name = "Distractor Type", values = vmacCols) +
  scale_colour_manual(name = "Distractor Type", values = vmacCols) +
  scale_shape_manual(name = "Distractor Type", guide = guide_legend(), values = c(21,22,23)) +
  ylab("% First Saccades") + xlab("Stimulus Type") + labs(tag = "C", title = "Feature Search") +
  my_theme

legend_saccade <- get_legend(saccadePlot_singleton + theme(legend.position = "bottom",
                                                           legend.title = element_text(size = 14),
                                                           legend.text = element_text(size = 14),
                                                           legend.key.width = unit(24, 'pt')))

## Capture Bar Plots
captureBarPlot_singleton <- captureMeans %>%
  filter(condition == "Singleton Search") %>%
  mutate(annotation = c("***", " ")) %>%
  ggplot(aes(x = dType, y = Capture, fill = dType)) +
  annotate(geom = "text", label = "Capture", x = 2.85, y = .15, angle = 90, vjust = 1, hjust = .5, family = font_rc) +
  annotate(geom = "text", label = "Suppression", x = 2.85, y = -.15, angle = 90, vjust = 1, hjust = .5, family = font_rc) +
  annotate(geom = "segment", x = 2.65, xend = 2.65, y = -.03, yend = -.27, arrow = arrow(type = 'closed', length = unit(6, 'pt'))) +
  annotate(geom = "segment", x = 2.65, xend = 2.65, y = .03, yend = .27, arrow = arrow(type = 'closed', length = unit(6, 'pt'))) +
  geom_col(position = position_dodge(.9), size = .3) +
  geom_hline(yintercept = 0, color = "black", size = .5) +
  geom_errorbar(aes(ymax = Capture + se, ymin = Capture - se), position = position_dodge(.9), width = .1) +
  geom_text(aes(label = annotation), nudge_y = .05) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = c(0,0), limits = c(-.3001, .3001)) +
  scale_x_discrete(labels = c("High\nreward", "Low\nreward")) +
  coord_cartesian(xlim = c(1,2),
                  clip = 'off') +
  scale_fill_manual(name = "Distractor Type", values = vmacCols, guide = guide_legend()) +
  ylab("Capture Score") + xlab("Trial Type") + labs(tag = "B") +
  my_theme

captureBarPlot_feature <- 
  captureMeans %>%
  filter(condition == "Feature Search") %>%
  mutate(annotation = c("***", "*")) %>%
  ggplot(aes(x = dType, y = Capture, fill = dType)) +
  annotate(geom = "text", label = "Capture", x = 2.85, y = .15, angle = 90, vjust = 1, hjust = .5, family = font_rc) +
  annotate(geom = "text", label = "Suppression", x = 2.85, y = -.15, angle = 90, vjust = 1, hjust = .5, family = font_rc) +
  annotate(geom = "segment", x = 2.65, xend = 2.65, y = -.03, yend = -.27, arrow = arrow(type = 'closed', length = unit(6, 'pt'))) +
  annotate(geom = "segment", x = 2.65, xend = 2.65, y = .03, yend = .27, arrow = arrow(type = 'closed', length = unit(6, 'pt'))) +
  geom_col(position = position_dodge(.9), size = .3) + 
  geom_hline(yintercept = 0, size = .5) +
  geom_errorbar(aes(ymax = Capture + se, ymin = Capture - se), position = position_dodge(.9), width = .1) +
  geom_text(aes(label = annotation), nudge_y = c(.05, .07)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = c(0,0), limits = c(-.3001, .3001)) +
  scale_x_discrete(labels = c("High\nreward", "Low\nreward")) +
  coord_cartesian(xlim = c(1,2),
                  clip = 'off') +
  scale_fill_manual(name = "Distractor Type", values = vmacCols, guide = guide_legend()) +
  ylab("Capture Score") + xlab("Trial Type") + labs(tag = "D") +
  my_theme

saccadePlot_ms_row1 <- saccadePlot_singleton + captureBarPlot_singleton + plot_spacer() + plot_layout(nrow = 1, widths = c(2,1,.1))

saccadePlot_ms_row2 <- saccadePlot_feature + captureBarPlot_feature + plot_spacer() + plot_layout(nrow = 1, widths = c(2,1,.1))

saccadePlot_ms <- saccadePlot_ms_row1 / legend_saccade / saccadePlot_ms_row2 + plot_layout(heights = c(1,.1,1))

save_plot(here("analysis", "figures", "saccade_plot.png"), saccadePlot_ms, ncol = 2, nrow = 2, type = "cairo", base_aspect_ratio = .9, dpi = 300)

saccadePlot_ms
```

```{r First Saccade Direction - Capture ANOVA}
capture.ANOVA <- aov_car(Capture ~ condition*dType + Error(sub/dType), data = captureData_means_byparticipant,  anova_table = list(es="pes"))

capture.ANOVA

# Bayes Factor for interaction contrast
captureData_wide <- captureData_means_byparticipant %>%
  select(sub, condition, dType, Capture) %>%
  spread(dType, Capture) %>%
  mutate(vmac = High - Low)

capture_bayes <- ttestBF(x = captureData_wide$vmac[captureData_wide$condition == "Singleton Search"], 
                         y = captureData_wide$vmac[captureData_wide$condition == "Feature Search"])

1/capture_bayes

```{r First Saccade Direction - Planned t-tests}
Singleton_high.ttest <- captureData_means_byparticipant %>%
  ungroup() %>%
  filter(condition == "Singleton Search") %>%
  filter(dType == "High") %>%
  select(Capture) %>%
  pull() %>%
  t.test()

Singleton_high.es <- captureData_means_byparticipant %>%
  ungroup() %>%
  filter(condition == "Singleton Search") %>%
  filter(dType == "High") %>%
  select(Capture) %>%
  summarise(mean_capture = mean(Capture), sd_capture = sd(Capture), d_z = mean_capture/sd_capture) %>%
  pull(d_z)

Singleton_low.ttest <- captureData_means_byparticipant %>%
  ungroup() %>%
  filter(condition == "Singleton Search") %>%
  filter(dType == "Low") %>%
  select(Capture) %>%
  pull() %>%
  t.test()

Singleton_low.es <- captureData_means_byparticipant %>%
  ungroup() %>%
  filter(condition == "Singleton Search") %>%
  filter(dType == "Low") %>%
  select(Capture) %>%
  summarise(mean_capture = mean(Capture), sd_capture = sd(Capture), d_z = mean_capture/sd_capture) %>%
  pull(d_z)

Feature_high.ttest <- captureData_means_byparticipant %>%
  ungroup() %>%
  filter(condition == "Feature Search") %>%
  filter(dType == "High") %>%
  select(Capture) %>%
  pull() %>%
  t.test()

Feature_high.es <- captureData_means_byparticipant %>%
  ungroup() %>%
  filter(condition == "Feature Search") %>%
  filter(dType == "High") %>%
  select(Capture) %>%
  summarise(mean_capture = mean(Capture), sd_capture = sd(Capture), d_z = mean_capture/sd_capture) %>%
  pull(d_z)

Feature_low.ttest <- captureData_means_byparticipant %>%
  ungroup() %>%
  filter(condition == "Feature Search") %>%
  filter(dType == "Low") %>%
  select(Capture) %>%
  pull() %>%
  t.test()

Feature_low.ttest$statistic <- Feature_low.ttest$statistic*-1 # just making the t value positive for reporting in the manuscript

Feature_low.es <- captureData_means_byparticipant %>%
  ungroup() %>%
  filter(condition == "Feature Search") %>%
  filter(dType == "Low") %>%
  select(Capture) %>%
  summarise(mean_capture = mean(Capture), sd_capture = sd(Capture), d_z = abs(mean_capture/sd_capture)) %>%
  pull(d_z)
```
